\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-09 paper/draft.tex
end-to-end pass

% TODO Add Ignis and Mozart
%% \documentclass[letterpaper,twocolumn,10pt]{article}
%% \pdfoutput=1 % for arxiv
%% \usepackage{usenix2019_v3}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
This paper presents \sys, a system aimed at parallelizing POSIX shell scripts through a combination of program transformations and runtime primitives.
% a source-to-source shell compiler and associated runtime library
Given a script, \sys converts it to a data-flow graph, performs a series of analyses that aim at exposing parallelism while preserving its sequential semantics, and then converts the data-flow graph back into a script that uses POSIX constructs to explicitly guide parallelism.
A lightweight annotation language allow command developers to express key parallelizability properties about their commands.
An accompanying parallelizability study of POSIX and GNU commands, two large and commonly used groups, 
informs \sys's annotation language and guides the development of a data-parallel standard library.
% \kk{eurosys: Can we extend this to say that PaSh automatically uses constructs from this library? Otherwise it sounds like the user has to use it to make their scripts parallel.}
Finally, \sys's \unix-aware runtime primitives address several practical issues related to performance and correctness.
\sys's extensive evaluation, combining several classic and modern \unix scripts, shows significant \todo{speedups} stemming from the combination of its program transformations and runtime primitives.
\end{abstract}

@@

\nv{Make sure we use \ttt{width} everywhere}
\nv{Decide on parallelizable regions}

@@

The second group is \emph{shell users}, who use POSIX shell constructs to combine multiple such commands from many languages into their scripts and are thus left with only a few options for incorporating parallelism.
% \kk{I don't really like the reference to such a later section. Can we remove it?}
% \nv{We have to say this is not just ``an opinion'', but that we give a scientific treatment/full coverage in the related work section. How far away should the sections be so that you're comfortable referencing to?}
% If versed into parallelism, 
% Users versed in computing can apply 
% these tools are command-unaware thus at risk of breaking program semantics,
% require manual changes to the code to make parallelism explicit,
% at times leverage command flags like the ones mentioned earlier---but each command 
% oblivious? agnostic?  
One option is manual tools such as GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{ts}~\cite{tsp}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm};
  these tools are either command-unaware, and thus at risk of breaking program semantics, or too coarse-grained, and thus only capable of exploiting parallelism at the level of entire scripts rather than individual components.
% Tools will attempt to parallelize anything 
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
% (ii) ad-hoc flags provided by command developers and mentioned above.
Another option is to use shell primitives (such as \ttt{&}, \ttt{wait}) to explicitly induce parallelism;
 these come at a cost of manual effort to split inputs, rewrite scripts, and orchestrate execution---an expensive and error-prone process.
% re-writing of the entire script from scratch; unfortunately, this process is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
To top it off, all these options assume a good understanding of parallelism;
  users with domain of expertise outside computing---hobbyists to data analysts---are left without options.
%  (see~\S\ref{related} for a full discussion).
% \nv{I still don't like ``for a full discussion''---raises too many expectations, and we don't have that much more to say}

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/pash_overview.pdf}}
\caption{
  \textbf{\sys overview and correspondence to sections.}
  \sys identifies dataflow regions~\sx{dataflow-regions}, 
  converts them to a dataflow graph~\sx{graph-components}, applies iterative
  transformations~\sx{ir:transformations} based on the parallelizability properties (\S\ref{cmd}, \S\ref{ext}) 
  of the commands used in these regions, and finally emits a parallel script
  that uses custom primitives~\sx{optimizer}.
}
\vspace{-15pt}
\label{fig:overview}
\end{figure}

@@

The above are tied together with \sys's runtime component.
% \sys's runtime  component additional packs
% Aside from these two components, \sys packs a runtime component that serves dual purpose:
Aware of the \unix philosophy and abstractions, it packs a small library of highly-optimized data aggregators as well as high-performance primitives for eager data splitting and merging.
These address many practical challenges % such as task parallelism, deadlocks, and performance bottlenecks,
  and were developed by uncovering several pathological situations, on a few of which we report.

@@

% \sys packs several ideas around the insight that a large subset of the Shell language can be represented as a data-flow graph, enabling for parallelization optimizations.
% First, a careful study of the parallelizability properties of shell primitives and commands.
% \sys addresses the challenge by offering command developers a unified set of abstractions for describing the parallelizability characteristics of their commands.
% Using lightweight annotations akin to type signatures, \sys can automate the parallelization of entire POSIX scripts without any input from their user.
% hand-in-hand with the parallelizability study:
%   the study informed the annotation language and was used to annotate the entire
%   for all 
% ---a set of annotations that \sys already offers for 
% for the standard library
%   akin to the standard library
%   the annotation language is informed by the study, but also the study is 
% XX to a type signature, these annotations describe 
% 
% Second, \sys defines a data-flow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the data-flow graph.
% (one sentence for the model)
% (one sentence for the optimizations)
% The result of the compilation is still a POSIX-compliant script that manages parallelism explicitly;
%   POSIX compliance means 
% it does not depend on a managed runtime 
% 
% Third, \sys packs a runtime component 
% A small library of highly-optimized data combiners allows 
% This is paired with a
% it comes with a few runtime primitives for eager data splitting and merging that address several practical issues ranging from deadlocks, performance bottlenecks;
%   these primitives were developed from uncovering several pathological situations, on many of which we report.

@@

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic \unix one-liners to modern data-processing scripts;
  (ii) two large and complex use cases for temperature analysis and web indexing;
%% , and container orchestration;
and 
  (iii) a series of micro-benchmarks inspired by (and
comparing with) possible alternatives.  The results show
significant advantages in correctness, manual effort,
and performance---with 
speedups $X-X\times$ due to \sys's transformations and $Y-Y\times$ due to \sys's runtime primitives.
%% ---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
Improvements depend on several factors such as a script's initial execution time and the parallelizability properties of its commands.

@@

\heading{Implementation} %
On top of command and shell semantics, the \unix environment has its own set of quirks.
Any attempt to orchestrate parallel execution will hit challenges related to task parallelism, deadlock prevention, and runtime performance.
For example, forked processes piping their combined results to Fig.~\ref{fig:example}'s \ttt{head} might not receive a \ttt{PIPE} signal if \ttt{head} exits prior to opening all pipes.
% A final challenge has to do with practical issues, as even manual parallelization of a given shell script is far from trivial.
% Parallelizing a script without reimplementing all of its commands, requires primitives to compose existing commands to achieve parallelization.
%  Designing correct and efficient compositions is particularly challenging in the context of \unix, and requires addressing issues related to scheduling, synchronization, and deadlock prevention. 
Moreover, several commands such as \ttt{sort} and \ttt{uniq} require specialized data aggregators in order to be correctly parallelized.
%% Developing correct, high-performance combiners for certain classes of commands such as \ttt{sort} and \ttt{uniq} is non trivial.
%% Moreover, the \unix environment has its set of quirks % about opening streams, deadling with FIFOs,
%%   and any attempt at parallelization will have to take care of these challenges---related to task parallelism, deadlock prevention, and runtime performance.

@@

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and a parameter \ttt{--width} and outputs a new script, a fragment of which is shown in Fig.~\ref{fig:example2}.
% concurrent version
% To expose and exploit data parallelism, % in a given shell script,
% %% parallel execution,
% \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% 
% \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{ir} by identifying and transforming special regions termed \emph{parallelizable regions} explained below. % nv idea: let's call them may-par regions
% It then runs a series of iterative analyses and transformations within these regions aimed at pushing parallelism as far as possible without risking breaking semantics.
% Finally, the compiler converts the DFG back into a shell script, replacing nodes with instantiated commands or primitives taken from its runtime library~\sx{optimizer}.
% An overview of \sys is shown in Fig.~\ref{fig:overview} and described below.
% First, \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model % ~\sx{dataflow-regions}
%   by identifying and transforming special regions termed \emph{parallelizable regions} and explained below. % nv idea: let's call them may-par regions
% \kk{eurosys: I suggest to replace the above sentence with the one below. There are a couple reasons. First I don't like the term source-to-source compiler because this sentence only describes the shell-graph compilation pass. Second, I would like to have some intuition about what there regions are (rather than deferring explanation for later). Third each region is tranformed to a DFG and not the whole script.}
\sys first identifies sections of the script that lack synchronization and scheduling constraints and converts them to dataflow graphs (DFGs).
It then performs a series of DFG transformations that expose parallelism without breaking semantics, by expanding the DFG to the desired \ttt{width}.
. % ~\sx{ir:transformations}.
Finally, \sys converts these DFGs back to a shell script augmented with \sys-provided commands. % ~\sx{optimizer}.
The script is handed off to the user's original shell interpreter for execution.
\sys addresses the aforementioned challenges~\sx{bg:challenges} as below.

@@

\begin{figure}[t]
\centering
% \includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_example2.pdf}}
\begin{minted}[fontsize=\footnotesize]{bash}
mkfifo $t{0,1...}
curl $base/$y > $t0 & cat $t0 | split $t1 $t2 &
cat $t1 | xargs -n1 curl -s >$t3 & 
cat $t2 | xargs -n1 curl -s >$t4 &
...
cat $t9 | sort -rn > $t11 & cat $t10 | sort -rn > $t12 &
cat $t11 | eager > $t13 & cat $t12 | eager > $t14 &
sort-merge -rn $t13 $t14 > $t15 &
cat $t15 | head -n1 > $out1 &
wait $! && ./pash_get_pids | xargs -n 1 kill -SIGPIPE
\end{minted}
\caption{
  \textbf{\sys's $2\times$-parallel output for fragment of Fig.~\ref{fig:example}.}
  \sys orchestrates the script's parallel execution through named pipes, parallel operators and careful placement of combiners.
}
\vspace{-15pt}
\label{fig:example2}
\end{figure}

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} containing commands that are candidates for parallelization~\sx{dataflow-regions}.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs expose task parallelism; others enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DFG)~\sx{graph-components}, 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent~\sx{ir:transformations}.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program~\sx{backend}.

@@

Broadly speaking, shell commands can be split into four major classes with respect to their parallelization characteristics, depending on what kind of state they need when processing their input (Tab.\ref{tab:classes}).
%% how they interact with state.
These classes are ordered in ascending difficulty (or impossibility) of parallelization.
In this order, some classes can be thought as subsets of the next---\eg all stateless commands are pure---meaning that the synchronization mechanisms required for any superclass would work with its subclass (but foregoing any performance improvements).
Commands can change classes depending on their flags,
% we focus on no-flag defaults for now and 
which are discussed later~\sx{ext}.

@@

\heading{Stateless Commands}
The first class, \sta, contains commands that operate on individual line elements of their input, without maintaining state across invocations.
These are commands that can be expressed as a purely functional $map$ or $filter$---\eg \ttt{grep} filters out individual lines and \ttt{basename} removes a path prefix from a string.
They may produce multiple elements---\eg \ttt{tr} may insert {\sc NL} tokens---but return empty output for empty input.
Workloads that use only stateless commands are trivial to parallelize:
  they do not require any synchronization to maintain correctness, nor caution about where to split inputs.
 % map :: (a -> b) -> [a] -> [b]

The choice of line as the data element strikes a convenient balance between coarse-grained (files) and fine-grained (characters) separation while staying in line with \unix's core abstractions.
This choice can affect the allocation of commands in \sta, as many of its commands (about 1/3) are stateless \emph{within} a stream element---\eg\ttt{tr} transliterates characters within a line, one at a time---enabling further parallelization by splitting individual lines.
This feature may seem of limited use, as these commands are computationally inexpensive, precisely due to their narrow focus.
However, it turns out to be  useful for cases with very large stream elements (\ie long lines) such the \ttt{.fastq} format used in bioinformatics.

@@

\heading{Key Concerns}
\sys's annotations focus on three crucial concerns about a command:
  (i) its parallelizability class,
  (ii) its inputs and outputs, and the characteristics of its input consumption, and
  % (ii) for commands that support multiple inputs, the characteristics of input consumption, and
  (iii) how flags affect its class, inputs, and outputs.
The first concern was discussed extensively in the previous section; we now turn to the latter two.

To be able to accurately construct the DFG representation, \sys needs to know certain details about a command's inputs and outputs.
There are a few reasons for this.
One reason is due to \sys's DFG representation, which connects commands with each other;
  to perform this connection correctly \sys needs to know the outputs of a command to correctly connect them with the inputs of the next.
A second reason is due to ordering:
  as some commands consume their inputs in a certain order, this order needs to be maintained in the parallel version of the program.
For example, consider the data-parallel version of \ttt{grep "foo" f1 f2}:

@@

Command flags (options) are a particularly popular way to control a command's execution, directly affecting its parallelizability classification.
Commands are thus assigned a default parallelizability class, which should then be refined by the set of flags the command uses.
For example, \ttt{cat} defaults to \sta, but with \ttt{-n} it jumps into \pur because it has to keep track of a counter and print it along with each line.

@@

\heading{Example Annotations}
The parallelizability properties of each command are specified in an annotation record that contains several components, each addressing one of the key concerns.
On the top level, each annotation record contains a set of clauses, each one identified by a predicate on the command arguments (Concern (iii)). 
Each clause contains an assignment for the command, indicating its class (Concern (i)), and the sequence of its inputs and outputs (Concern (ii)).
%  (a JSON object) that
% Annotations treat flag arguments differently from file arguments, by checking if they start with a \ttt{-}.
% \footnote{
%   Note that the single character \ttt{-} is not an option as it is most often used to designate that a command reads from its standard input.
%   \nv{
%     There are options of of the form --f "x; y", and BSD options do not take \ttt{-} (like \ttt{ps aux}).
%     I am also not sure how important this last sentence is, as it did not answer a question I had---I expected arguments to be anything: if soneome wants to put ``pizza'' the DSL should not disallow it.
%   }
% }
% The complete annotation grammar is shown in Appendix~\ref{alg}. \str

For example, consider \ttt{comm}, which performs a join-like operation on its two inputs to identify common elements.
When \ttt{comm} is invoked without any flags, it produces a three-column output:
  the first column contains lines that are unique to the first input, the second column contains lines that are unique to the second input, and the last column contains lines that exist in both inputs.
A user can invoke \ttt{comm} with any combination of flags \ttt{-1}, \ttt{-2}, or \ttt{-3} to suppress the corresponding output column(s).
% The first clause from its annotation record is shown below:
Below is its annotation record (left) and the implementation of record's first clause (right):
%
% \begin{minted}[fontsize=\footnotesize]{json-object}
%   "predicate":
%     { 
%       "operator": "all",
%       "operands": ["-1", "-3"] 
%     },    
%   "class": "stateless",
%   "inputs": ["args[1]"],
%   "outputs": ["stdout"],
%   "options": ["stdin-hyphen"]
% \end{minted}

>>>>

% TODO Add Ignis and Mozart
% \nv{Make sure we use \ttt{width} everywhere}
% \nv{Decide on parallelizable regions}

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
This paper presents \sys, a system aimed at parallelizing POSIX shell scripts through a combination of program transformations and runtime primitives.
% a source-to-source shell compiler and associated runtime library
Given a script, \sys converts it to a dataflow graph, performs a series of analyses and transformations that aim at exposing parallelism while preserving its sequential semantics, and then converts the dataflow graph back into a script that uses POSIX constructs to explicitly guide parallelism.
A lightweight annotation language allow command developers to express key parallelizability properties about their commands.
An accompanying parallelizability study of POSIX and GNU commands---two large and commonly used groups---informs \sys's annotation language and guides the development of a data-parallel standard library.
% \kk{eurosys: Can we extend this to say that PaSh automatically uses constructs from this library? Otherwise it sounds like the user has to use it to make their scripts parallel.}
Finally, \sys's \unix-aware runtime primitives address several practical issues related to performance and correctness.
\sys's extensive evaluation, combining several classic and modern \unix scripts, shows significant \todo{speedups} stemming from the combination of its program transformations and runtime primitives.
\end{abstract}

@@

@@

The second group is \emph{shell users}, who use POSIX shell constructs to combine multiple such commands from many languages into their scripts and are thus left with only a few options for incorporating parallelism.
% \kk{I don't really like the reference to such a later section. Can we remove it?}
% \nv{We have to say this is not just ``an opinion'', but that we give a scientific treatment/full coverage in the related work section. How far away should the sections be so that you're comfortable referencing to?}
% If versed into parallelism, 
% Users versed in computing can apply 
% these tools are command-unaware thus at risk of breaking program semantics,
% require manual changes to the code to make parallelism explicit,
% at times leverage command flags like the ones mentioned earlier---but each command 
% oblivious? agnostic?  
One option is to use manual tools such as GNU \ttt{parallel}~\cite{Tange2011a}, \ttt{ts}~\cite{tsp}, \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm};
  these tools are either command-unaware, and thus at risk of breaking program semantics, or too coarse-grained, and thus only capable of exploiting parallelism at the level of entire scripts rather than individual components.
% Tools will attempt to parallelize anything 
% use the composition abstractions of the \unix shell, 
% and are often prototypical in nature---focusing on quickly prototyping a solution for the task at hand rather building an optimal solution to a recurring pattern.
% short lived, changing quickly
% (ii) ad-hoc flags provided by command developers and mentioned above.
Another option is to use shell primitives (such as \ttt{&}, \ttt{wait}) to explicitly induce parallelism;
 these come at a cost of manual effort to split inputs, rewrite scripts, and orchestrate execution---an expensive and error-prone process.
% re-writing of the entire script from scratch; unfortunately, this process is expensive and can introduce new bugs, cascading changes, and divergence from legacy functionality---but also foregoes all the productivity benefits of the original script.
% The fact that a programmer often belongs to both groups at the same time---as composed scripts can themselves form a command used in other scrips---compounds the challenge.
To top it off, all these options assume a good understanding of parallelism;
  users with domain of expertise outside computing---from hobbyists to data analysts---are left without options.
%  (see~\S\ref{related} for a full discussion).
% \nv{I still don't like ``for a full discussion''---raises too many expectations, and we don't have that much more to say}

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{\detokenize{./figs/pash_overview.pdf}}
\caption{
  \textbf{\sys overview and correspondence to sections.}
  \sys identifies dataflow regions~\sx{dataflow-regions}, 
  converts them to a dataflow graph~\sx{graph-components}, applies iterative
  transformations~\sx{ir:transformations} based on the parallelizability properties (\S\ref{cmd}, \S\ref{ext}) 
  of the commands used in these regions, and finally emits a parallel script
  that uses custom primitives~\sx{optimizer}.
}
\vspace{-25pt}
\label{fig:overview}
\end{figure}

@@

The above are tied together with \sys's runtime component.
% \sys's runtime  component additional packs
% Aside from these two components, \sys packs a runtime component that serves dual purpose:
Aware of the \unix philosophy and abstractions, it packs a small library of highly-optimized data aggregators as well as high-performance primitives such as ones for eager data splitting and merging.
These address many practical challenges % such as task parallelism, deadlocks, and performance bottlenecks,
  and were developed by uncovering several pathological situations, on a few of which we report.

@@

% \sys packs several ideas around the insight that a large subset of the Shell language can be represented as a dataflow graph, enabling for parallelization optimizations.
% First, a careful study of the parallelizability properties of shell primitives and commands.
% \sys addresses the challenge by offering command developers a unified set of abstractions for describing the parallelizability characteristics of their commands.
% Using lightweight annotations akin to type signatures, \sys can automate the parallelization of entire POSIX scripts without any input from their user.
% hand-in-hand with the parallelizability study:
%   the study informed the annotation language and was used to annotate the entire
%   for all 
% ---a set of annotations that \sys already offers for 
% for the standard library
%   akin to the standard library
%   the annotation language is informed by the study, but also the study is 
% XX to a type signature, these annotations describe 
% 
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
% (one sentence for the model)
% (one sentence for the optimizations)
% The result of the compilation is still a POSIX-compliant script that manages parallelism explicitly;
%   POSIX compliance means 
% it does not depend on a managed runtime 
% 
% Third, \sys packs a runtime component 
% A small library of highly-optimized data combiners allows 
% This is paired with a
% it comes with a few runtime primitives for eager data splitting and merging that address several practical issues ranging from deadlocks, performance bottlenecks;
%   these primitives were developed from uncovering several pathological situations, on many of which we report.

@@

We evaluate \sys using
  (i) a series of benchmarks, ranging from classic \unix one-liners to modern data-processing scripts;
  (ii) two large and complex use cases for temperature analysis and web indexing;
%% , and container orchestration;
and 
  (iii) a series of micro-benchmarks inspired by (and comparing with) possible alternatives.
The vast majority (47/52) of scripts---all unmodified---see significant speedups.
Indicatively, applied to 10 programs with parallelism \ttt{--width=64}, \sys achieves 9.35$\times$ due to program transformations alone and 13.24$\times$ when coupled with its runtime primitives.
% \sys performs better than \ttt{sort --parallel}, a hand-tuned parallel implementation, and GNU parallel, which returns incorrect results if used without care.
In another case of \ttt{--width=16}, \sys successfully accelerates a large program by 2.44$\times$---extracting 10.79$\times$ from the portion of the program that traditionally falls outside the focus of conventional parallelization systems.

% For the temperature analysis script (which is taken from a Hadoop case study), \sys not only 
% contains pre-processing that takes X time and Y main processing (Hadoop focus).
% \sys accelerates the entire pipeline by , including the fragment that is outside the focus of conventional parallelization systems.
% accelerates the CPU-intensive part that is parallelized by Hadoop, but also for the preprocessing part which takes even more time than what the other part takes.
% XX--YY speedups in the vast majority of scripts, ones taken from the wild 
% 1. Sentence about transformations vs runtime primitives
% 2. Unix 50? 
% 3. point about Hadoop expressing a fraction of the pipeline and speedup
% \sys achieves higher speedup than \ttt{sort --parallel} that involved hand-tuned parallel implementation by command developers.

%   The results show
% significant advantages in correctness, manual effort,
% and performance---with 
% speedups $X-X\times$ due to \sys's transformations and $Y-Y\times$ due to \sys's runtime primitives.
%% ---with \sys's transformations offering an order-of-magnitude performance improvements aided further by \sys's optimized runtime components.
% Improvements depend on several factors such as a script's initial execution time and the parallelizability properties of its commands.

@@

\heading{Implementation} %
On top of command and shell semantics, the broader \unix environment has its own set of quirks.
Any attempt to orchestrate parallel execution will hit challenges related to task parallelism, deadlock prevention, and runtime performance.
For example, forked processes piping their combined results to Fig.~\ref{fig:example}'s \ttt{head} may not receive a \ttt{PIPE} signal if \ttt{head} exits prior to opening all pipes.
% A final challenge has to do with practical issues, as even manual parallelization of a given shell script is far from trivial.
% Parallelizing a script without reimplementing all of its commands, requires primitives to compose existing commands to achieve parallelization.
%  Designing correct and efficient compositions is particularly challenging in the context of \unix, and requires addressing issues related to scheduling, synchronization, and deadlock prevention. 
Moreover, several commands such as \ttt{sort} and \ttt{uniq} require specialized data aggregators in order to be correctly parallelized.
%% Developing correct, high-performance combiners for certain classes of commands such as \ttt{sort} and \ttt{uniq} is non trivial.
%% Moreover, the \unix environment has its set of quirks % about opening streams, deadling with FIFOs,
%%   and any attempt at parallelization will have to take care of these challenges---related to task parallelism, deadlock prevention, and runtime performance.

@@

At a high level, \sys takes as input a POSIX shell script like the one in Fig.~\ref{fig:example} and a parameter \ttt{--width} and outputs a new script, a fragment of which is shown in Fig.~\ref{fig:example2}.
% concurrent version
% To expose and exploit data parallelism, % in a given shell script,
% %% parallel execution,
% \sys analyzes individual program fragments, addressing the challenges mentioned earlier.
% 
% \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model~\sx{ir} by identifying and transforming special regions termed \emph{dataflow regions} explained below. % nv idea: let's call them may-par regions
% It then runs a series of iterative analyses and transformations within these regions aimed at pushing parallelism as far as possible without risking breaking semantics.
% Finally, the compiler converts the DFG back into a shell script, replacing nodes with instantiated commands or primitives taken from its runtime library~\sx{optimizer}.
% An overview of \sys is shown in Fig.~\ref{fig:overview} and described below.
% First, \sys's source-to-source compiler translates the script to a dataflow graph (DFG) model % ~\sx{dataflow-regions}
%   by identifying and transforming special regions termed \emph{dataflow regions} and explained below. % nv idea: let's call them may-par regions
% \kk{eurosys: I suggest to replace the above sentence with the one below. There are a couple reasons. First I don't like the term source-to-source compiler because this sentence only describes the shell-graph compilation pass. Second, I would like to have some intuition about what there regions are (rather than deferring explanation for later). Third each region is tranformed to a DFG and not the whole script.}
\sys first identifies sections of the script that lack synchronization and scheduling constraints and converts them to dataflow graphs (DFGs).
It then performs a series of DFG transformations that expose parallelism without breaking semantics, by expanding the DFG to the desired \ttt{width}.
. % ~\sx{ir:transformations}.
Finally, \sys converts these DFGs back to a shell script augmented with \sys-provided commands. % ~\sx{optimizer}.
The script is handed off to the user's original shell interpreter for execution.
\sys addresses the aforementioned challenges~\sx{bg:challenges} as below.

@@

\begin{figure}[t]
\centering
% \includegraphics[width=0.49\textwidth]{\detokenize{./figs/dish_example2.pdf}}
\begin{minted}[fontsize=\footnotesize,escapeinside=()]{bash}
mkfifo $t{0,1...}
curl $base/$y > $t0 & cat $t0 | split $t1 $t2 &
cat $t1 | xargs -n1 curl -s >$t3 & 
cat $t2 | xargs -n1 curl -s >$t4 &
...
cat $t9 | sort -rn > $t11 & cat $t10 | sort -rn > $t12 &
cat $t11 | eager > $t13 & cat $t12 | eager > $t14 &
sort-merge -rn $t13 $t14 > $t15 &
cat $t15 | head -n1 > $out1 &
wait $! && get-pids | xargs -n 1 kill -SIGPIPE
\end{minted}
\caption{
  \textbf{Output of \ttiny{pash --width=2} for Fig.~\ref{fig:example} (fragment).}
  \sys orchestrates parallel execution through named pipes, parallel operators, and custom runtime primitives---\eg \ttiny{eager}, \ttiny{sort-merge}, and \ttiny{get-pids}.
}
\vspace{-15pt}
\label{fig:example2}
\end{figure}

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{dataflow regions} containing commands that are candidates for parallelization~\sx{dataflow-regions}.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs expose parallelism; others enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each dataflow region to a \emph{dataflow graph} (DFG)~\sx{graph-components}, 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a dataflow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent~\sx{ir:transformations}.
%% , in which DFGs indicate the dataflow regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program~\sx{backend}.

@@

Broadly speaking, shell commands can be split into four major classes with respect to their parallelization characteristics, depending on what kind of state they mutate when processing their input (Tab.\ref{tab:classes}).
%% how they interact with state.
These classes are ordered in ascending difficulty (or impossibility) of parallelization.
In this order, some classes can be thought as subsets of the next---\eg all stateless commands are pure---meaning that the synchronization mechanisms required for any superclass would work with its subclass (but foregoing any performance improvements).
Commands can change classes depending on their flags,
% we focus on no-flag defaults for now and 
which are discussed later~\sx{ext}.

@@

\heading{Stateless Commands}
The first class, \sta, contains commands that operate on individual line elements of their input, without maintaining state across invocations.
These are commands that can be expressed as a purely functional $map$ or $filter$---\eg \ttt{grep} filters out individual lines and \ttt{basename} removes a path prefix from a string.
They may produce multiple elements---\eg \ttt{tr} may insert {\sc NL} tokens---but always return empty output for empty input.
Workloads that use only stateless commands are trivial to parallelize:
  they do not require any synchronization to maintain correctness, nor caution about where to split inputs.
 % map :: (a -> b) -> [a] -> [b]

The choice of line as the data element strikes a convenient balance between coarse-grained (files) and fine-grained (characters) separation while staying in line with \unix's core abstractions.
This choice can affect the allocation of commands in \sta, as many of its commands (about 1/3) are stateless \emph{within} a stream element---\eg\ttt{tr} transliterates characters within a line, one at a time---enabling further parallelization by splitting individual lines.
This feature may seem of limited use, as these commands are computationally inexpensive, precisely due to their narrow focus.
However, it may be  useful for cases with very large stream elements (\ie long lines) such the \ttt{.fastq} format used in bioinformatics.

@@

\heading{Key Concerns}
\sys's annotations focus on three crucial concerns about a command:
  (C1) its parallelizability class,
  (C2) its inputs and outputs, and the characteristics of its input consumption, and
  % (ii) for commands that support multiple inputs, the characteristics of input consumption, and
  (C3) how flags affect its class, inputs, and outputs.
The first concern was discussed extensively in the previous section; we now turn to the latter two.

To be able to accurately construct the DFG representation, \sys needs to know certain details about a command's inputs and outputs.
There are a few reasons for this.
One reason is due to \sys's DFG representation, which connects commands with each other.
  To perform this connection correctly, \sys needs to know the outputs of a command. % to correctly connect them with the inputs of the next.
A second reason is due to ordering:
  as some commands consume their inputs in a certain order, this order needs to be maintained in the parallel version of the program.
For example, consider the data-parallel version of \ttt{grep "foo" f1 f2}:

@@

Command flags (options) are a particularly popular way of controlling a command's execution, directly affecting its parallelizability classification.
Commands are thus assigned a default parallelizability class, which is then refined by the set of flags the command uses.
For example, \ttt{cat} defaults to \sta, but with \ttt{-n} it jumps into \pur because it has to keep track of a counter and print it along with each line.

@@

\heading{Example Annotations}
The parallelizability properties of each command are specified in an annotation record that contains several components, each addressing one of the key concerns C1--3.
At the top level, each annotation record contains a set of clauses, each one identified by a predicate on the command arguments (C3). 
Each clause contains an assignment for the command, indicating its class (C1), and the sequence of its inputs and outputs (C2).
%  (a JSON object) that
% Annotations treat flag arguments differently from file arguments, by checking if they start with a \ttt{-}.
% \footnote{
%   Note that the single character \ttt{-} is not an option as it is most often used to designate that a command reads from its standard input.
%   \nv{
%     There are options of of the form --f "x; y", and BSD options do not take \ttt{-} (like \ttt{ps aux}).
%     I am also not sure how important this last sentence is, as it did not answer a question I had---I expected arguments to be anything: if soneome wants to put ``pizza'' the DSL should not disallow it.
%   }
% }
% The complete annotation grammar is shown in Appendix~\ref{alg}. \str

For example, consider \ttt{comm}, which performs a join-like operation on its two inputs to identify common elements.
When \ttt{comm} is invoked without any flags, it produces a three-column output:
  the first column contains lines that are unique to the first input, the second column contains lines that are unique to the second input, and the last column contains lines that exist in both inputs.
A user can invoke \ttt{comm} with any combination of flags \ttt{-1}, \ttt{-2}, or \ttt{-3} to suppress the corresponding output column(s).
% The first clause from its annotation record is shown below:
Below is its annotation record (left) and the implementation of the record's first clause (right):
%
% \begin{minted}[fontsize=\footnotesize]{json-object}
%   "predicate":
%     { 
%       "operator": "all",
%       "operands": ["-1", "-3"] 
%     },    
%   "class": "stateless",
%   "inputs": ["args[1]"],
%   "outputs": ["stdout"],
%   "options": ["stdin-hyphen"]
% \end{minted}
